### 2012-11-24 paper.tex
calculated default features -- introduced booktabs package

\usepackage{natbib}
\usepackage{listings}
\bibliographystyle{plain}

@@

\begin{lstlisting}
J48 pruned tree
------------------

survival <= 12
|   survival <= 11: 0 (4.0)
|   survival > 11: 1 (4.0)
survival > 12: 1 (87.0)
\end{lstlisting}

@@

The J48 Decision Tree shows a more useful tree on the re-preprocessed data:
\begin{lstlisting}
J48 pruned tree

epss <= 13100: 1 (90.63/1.95)
epss > 13100
|   fractional-shortening <= 190: 1 (2.25/0.05)
|   fractional-shortening > 190: 0 (2.11/0.11)
\end{lstlisting}

@@

Using an unpruned tree doesn't improve the results:
\begin{lstlisting}
J48 unpruned tree

pericardial-effusion = 0
|   epss <= 13100: 1 (76.62/1.95)
|   epss > 13100
|   |   epss <= 15600: 0 (2.19/0.16)
|   |   epss > 15600: 1 (2.19/0.03)
pericardial-effusion = 1: 1 (14.0)
\end{lstlisting}
Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.

\subsubsection{Random Forest}
Let's see if a random forest in place of a simple decision tree can make better result out of the echocardiogram data set. With default settings, i.e. default number of features \todo[inline]{calculate!} and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:

\begin{lstlisting}
=== Confusion Matrix ===
  a  b   <-- classified as
  0  4 |  a = 0
  2 91 |  b = 1
\end{lstlisting}
And the statistical measures:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.022    0          0       0            0
                 0.978    1        0.958      0.978   0.968        1
Weighted Avg.    0.938    0.96     0.918      0.938   0.928     
\end{lstlisting}
\paragraph{Parameter: Number of trees}
\todo[inline]{proper line break here}
Reducing the number of the trees continuously decreases the quality of the model, as shown in the accuracy details for 5 random trees:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.032    0          0       0            0  
                 0.968    1        0.957      0.968   0.963        1   
Weighted Avg.    0.928    0.96     0.918      0.928   0.923    
\end{lstlisting}
.. and for a single random tree:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.054    0          0       0            0  
                 0.946    1        0.957      0.946   0.951        1   
Weighted Avg.    0.907    0.961    0.917      0.907   0.912    
\end{lstlisting}
Reducing the number of the trees increases the number of false positives but doesn't catch any true positives.
\paragraph{Limiting the tree-depth}
Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set. The only change to the outcome can be observed as soon the depth of the trees is limited to 6:
\begin{lstlisting}
    === Confusion Matrix ===
    a  b   <-- classified as
    0  4 |  a = 0
    1 92 |  b = 1
\end{lstlisting}
Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
\paragraph{Increasing the number of features}
Increasing the number of features chosen by each random tree seems to has also a low effect on the quality, because until a certain border of 5 chosen features the outcome is exactly the same as it was with default settings. But if we configure the random forests with 6 features we can spot our first true positive:
\begin{lstlisting}
=== Confusion Matrix ===
  a  b   <-- classified as
  1  3 |  a = 0
  2 91 |  b = 1
\end{lstlisting}
Which gives us the best achievable accuracy that we can reach with decision based algorithms:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure   Class
                0.25     0.022    0.333      0.25    0.286       0
                0.978    0.75     0.968      0.978   0.973       1
Weighted Avg.   0.948    0.72     0.942      0.948   0.945      
\end{lstlisting}

@@

With default settings, a random forest achieves an accuracy as follows:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
                0.615    0.108    0.644      0.615   0.629      >50K
                0.892    0.385    0.88       0.892   0.886      <=50K
Weighted Avg.   0.825    0.318    0.823      0.825   0.824     
\end{lstlisting}
which is surprisingly a worse result than from the J48 algorithm.

@@

\paragraph{Limitations of the maximum depth of the trees}
We have a sign for overfitting under default settings: Reducing the maximum depth has a positive effect on the quality. From size 20 over 10 until the size of 7 the precision of the random forest rises until 85\%. Reducing the depth even further lets the precision fall again. Here is the detailed accuracy for a random forest which can grow only to a height of 6:

@@

\paragraph{Number of features}
Increasing the number of features doesn't support the quality at all. The default number of features for this data set with 13 attributes is \todo[inline]{calculate!}. With a number of 5 randomly chosen features, the precision is yet at 85\% (with maximum depth of tree = 6) but with 10 randomly chosen features the precision falls back to 84.5\%.

>>>>

\usepackage{natbib}
\usepackage{listings}
\usepackage{booktabs}
\bibliographystyle{plain}

@@

\begin{lstlisting}
J48 pruned tree
------------------
survival <= 12
|   survival <= 11: 0 (4.0)
|   survival > 11: 1 (4.0)
survival > 12: 1 (87.0)
\end{lstlisting}

@@

The J48 Decision Tree shows a more useful tree on the re-preprocessed data:
\begin{lstlisting}
J48 pruned tree
epss <= 13100: 1 (90.63/1.95)
epss > 13100
|   fractional-shortening <= 190: 1 (2.25/0.05)
|   fractional-shortening > 190: 0 (2.11/0.11)
\end{lstlisting}

@@

Using an unpruned tree doesn't improve the results:
\begin{lstlisting}
J48 unpruned tree
pericardial-effusion = 0
|   epss <= 13100: 1 (76.62/1.95)
|   epss > 13100
|   |   epss <= 15600: 0 (2.19/0.16)
|   |   epss > 15600: 1 (2.19/0.03)
pericardial-effusion = 1: 1 (14.0)
\end{lstlisting}
Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.

\subsubsection{Random Forest}
Let's see if a random forest in place of a simple decision tree can make better result out of the echocardiogram data set. With default settings, i.e. default number of features $ log_2(7) + 1 \approx 4 $, and 10 random trees with unlimited tree depth, the random forest shows us a rather disappointing result:

\begin{lstlisting}
=== Confusion Matrix ===
  a  b   <-- classified as
  0  4 |  a = 0
  2 91 |  b = 1
\end{lstlisting}
And the statistical measures:
\begin{table}
    \begin{tabular}{lllllll}
        \toprule 
        \multicolumn{7}{c}{Detailed Accuracy By Class} \\ 
              &  TP Rate & FP Rate & Precision & Recall & F-Measure & Class \\ \midrule
              &  0       & 0.022   & 0         & 0      & 0         & 0     \\
              &  0.978   & 1       & 0.958     & 0.978  & 0.968     & 1     \\
Weighted Avg. &  0.938   & 0.96    & 0.918     & 0.938  & 0.928     &       \\ \bottomrule
    \end{tabular}
\end{table}

\paragraph{Number of Random Trees}
\todo[inline]{proper line break here}
Reducing the number of the trees continuously decreases the quality of the model, as shown in the accuracy details for 5 random trees:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.032    0          0       0            0  
                 0.968    1        0.957      0.968   0.963        1   
Weighted Avg.    0.928    0.96     0.918      0.928   0.923    
\end{lstlisting}
.. and for a single random tree:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure    Class
                 0        0.054    0          0       0            0  
                 0.946    1        0.957      0.946   0.951        1   
Weighted Avg.    0.907    0.961    0.917      0.907   0.912    
\end{lstlisting}
Reducing the number of the trees increases the number of false positives but doesn't catch any true positives.
\paragraph{Limiting the maximum depth of the trees}
Limitations in the depths of the random trees has also almost no effect on the accuracy of random forests on this data set. The only change to the outcome can be observed as soon the depth of the trees is limited to 6:
\begin{lstlisting}
    === Confusion Matrix ===
    a  b   <-- classified as
    0  4 |  a = 0
    1 92 |  b = 1
\end{lstlisting}
Only one false positive has gone. Not really a remarkable change in the quality. This does not change if the tree-depth is further reduced.
\paragraph{Increasing the number of features}
Increasing the number of features chosen by each random tree seems to has also a low effect on the quality, because until a certain border of 5 chosen features the outcome is exactly the same as it was with default settings. But if we configure the random forests with 6 features we can spot our first true positive:
\begin{lstlisting}
=== Confusion Matrix ===
  a  b   <-- classified as
  1  3 |  a = 0
  2 91 |  b = 1
\end{lstlisting}
Which gives us the best achievable accuracy that we can reach with decision based algorithms:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure   Class
                0.25     0.022    0.333      0.25    0.286       0
                0.978    0.75     0.968      0.978   0.973       1
Weighted Avg.   0.948    0.72     0.942      0.948   0.945      
\end{lstlisting}

@@

With default settings, a random forest achieves an accuracy as follows:
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                TP Rate  FP Rate  Precision  Recall  F-Measure  Class
                0.615    0.108    0.644      0.615   0.629      >50K
                0.892    0.385    0.88       0.892   0.886      <=50K
Weighted Avg.   0.825    0.318    0.823      0.825   0.824     
\end{lstlisting}
which is surprisingly worse than the J48 algorithm.

@@

\paragraph{Limiting the maximum depth of the trees}
We have a sign for overfitting under default settings: Reducing the maximum depth has a positive effect on the quality. From size 20 over 10 until the size of 7 the precision of the random forest rises until 85\%. Reducing the depth even further lets the precision fall again. Here is the detailed accuracy for a random forest which can grow only to a height of 6:

@@

\paragraph{Number of features}
Increasing the number of features doesn't support the quality at all. The default number of features for this data set with 13 attributes is $ log_2(13) + 1 \approx 5 $. Choosing 6 attributes brinchoose gs us a precision of 84,9\% and with 10 randomly chosen features the precision falls back to 84.5\%. We can tease the most out of the random forest if we decrease the number of features a little. To be concise, we get the best achievable result, if we take only 3 randomly chosen
features.
\begin{lstlisting}
=== Detailed Accuracy By Class ===
                 TP Rate  FP Rate  Precision  Recall  F-Measure Class
                 0.577    0.052    0.779      0.577   0.663     >50K
                 0.948    0.423    0.876      0.948   0.911     <=50K
Weighted Avg.    0.859    0.333    0.853      0.859   0.851    
\end{lstlisting}
Which is a quite similar accuracy like the best of the J48 algorithm. 
\subsection{Support Vector Machine}
